---
title: "Reproducible Research: Peer Assessment 1"
author: "Remco Keijser - JAN-2016"
keep_md: true
output: 
  html_document: 
    keep_md: yes
---


## Loading and preprocessing the data
```{r}
import.data <- read.csv(file="activity.csv", header=TRUE)
import.data$date <- as.Date(as.character(import.data$date, format = "%Y/%m/%d"))
```

```{r}
# INSTALL PACKAGES
# install.packages("dplyr")             # Install dplyr package
# install.packages("timeDate")          # Install timeDate package
# install.packages("lattice")           # Install lattice package

```

## What is mean total number of steps taken per day?
```{r, message=F, warning=F}
library("dplyr")
STATS <- summarise(group_by(import.data, date),
                   sum_NA = sum(is.na(steps)),
                   sum_steps = sum(steps, na.rm=TRUE),
                   mean_steps = mean(steps, na.rm=TRUE),
                   median_steps = median(steps, na.rm = TRUE)
)
```

```{r, fig.width=8, fig.height=6}
hist(STATS$sum_steps,
     main="Total number of steps taken each day\nWITH MISSING VALUES",
     xlab="Number of steps",
     border="black",
     col="grey",
     xlim=c(0,25000),
     las=1,
     breaks=5)
abline( v = mean(STATS$sum_steps), col=1, lty = 1)
abline( v = median(STATS$sum_steps), col=2, lty = 2)
legend("topright", legend = c("Mean", "Median"), col=c(1,2), lty=c(1,2))
```

```{r}
library(knitr)
STATS
```

## What is the average daily activity pattern?

```{R}
STATS2 <- summarise(group_by(import.data, interval),
                    average_steps = mean(steps, na.rm=TRUE))
```

```{r, fig.width=8, fig.height=6}
plot(STATS2$interval, STATS2$average_steps, type='l',
     main="Average daily activity pattern",
     xlab="Interval",
     ylab="Total Steps",
     col="blue",
     xlim=c(0,2500))
```

```{R}
max_steps <- max(STATS2$average_steps)
max_interval <- STATS2[STATS2$average_steps==max_steps,]
max_interval
```
#### **Interval 835 contains the maximum number of steps.**

## Imputing missing values

```{r}
library(knitr)
kable(summary(import.data))
```  

#### **In total 2304 missing values in variable steps. 8 days are missing step values**

```{r}
STATS[c(1,8,32,35,40,41,45,61),c(1,2)]
```

#### **Stratergy chosen to impute missing values: use the mean for that 5-minute interval.**

```{r}
import.data.imputed <- merge(import.data, STATS2, by.x = "interval", by.y = "interval")
import.data.imputed <- import.data.imputed[order(as.Date(import.data.imputed$date, format="%d/%m/%Y")),]
import.data.imputed$steps.imputed <- NA

for(i in 1:length(import.data.imputed$steps)){
        
        if(is.na(import.data.imputed[i,2])) {
                
                import.data.imputed[i,5] <- import.data.imputed[i,4]
        }
        
        if(!is.na(import.data.imputed[i,2])) {
                
                import.data.imputed[i,5] <- import.data.imputed[i,2]
        }
}

import.data.imputed <- import.data.imputed[,c(2,5,3,1)]
```

```{r}
library(knitr)
head(import.data.imputed, n=50)
```

```{r}
STATS3 <- summarise(group_by(import.data.imputed, date),
                    sum_NA = sum(is.na(steps)),
                    sum_NA.imputed = sum(is.na(steps.imputed)),
                    sum_steps = sum(steps, na.rm=TRUE),
                    mean_steps = mean(steps, na.rm=TRUE),
                    median_steps = median(steps, na.rm = TRUE),
                    
                    sum_steps.imputed = sum(steps.imputed, na.rm=TRUE),
                    mean_steps.imputed = mean(steps.imputed, na.rm=TRUE),
                    median_steps.imputed = median(steps.imputed, na.rm = TRUE))
```

```{r, fig.width=8, fig.height=6}
hist(STATS3$sum_steps.imputed,
     main="Total number of steps taken each day\nMISSING VALUES IMPUTED",
     xlab="Number of steps",
     border="black",
     col="grey",
     xlim=c(0,25000),
     las=1,
     breaks=5)
abline( v = mean(STATS3$sum_steps.imputed), col=1, lty = 1)
abline( v = median(STATS3$sum_steps.imputed), col=2, lty = 2)
legend("topright", legend = c("Mean", "Median"), col=c(1,2), lty=c(1,2))
```

```{r}
STATS3
```

#### **The mean and median differ of the imputed day.**
#### **As shown in the two histograms the mean and the median also changes slighly of the sum of the steps across al days**

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library("timeDate")
import.data.imputed$weekday <- isWeekday(import.data.imputed$date, wday = 1:5)
import.data.imputed$weekend <- isWeekend(import.data.imputed$date, wday = 1:5)
import.data.imputed$day <- import.data.imputed$weekday
import.data.imputed$day[import.data.imputed$day == TRUE] <- "WEEKDAY"
import.data.imputed$day[import.data.imputed$day == FALSE] <- "WEEKEND"
head(import.data.imputed, n=10)
```

```{r, fig.width=8, fig.height=6}
library("lattice")
xyplot(steps.imputed  ~ interval| day, data=import.data.imputed,
       type = "l",
       layout = c(1, 2),
       par.strip.text=list(cex=0.8),
       xlab = "Interval",
       ylab=list(
               label="Number of steps",
               cex=1),
       main=list(
               label="Interval vs number of steps",
               cex=1),
       scales=list(cex=0.8)
)
```
